Modeling Interventions

### 2020-06-06 model_interventions.Rmd
integrated CHILDES word frequency distribution into model, and now plot learning of some individual words and total vs. CDI words known

---
title: "Modeling Interventions"
output: html_notebook
---

@@

```{r}
require(tidyverse)
source("model-nonsampling.R")

# Gilkerson et al. 2017 Table 2: standardized AWC

@@

# Bergelson et al. 2018 Table S1: 
Bergelson_s1 <- tribble(
  ~study, ~AWC_mean, ~AWC_sd, ~CDS_mean, ~CDS_sd, ~N, 
  "Schneidman & Goldin-Meadow 2013", 
  "Schneidman & Goldin-Meadow 2013", 
  "Weisleder & Fernald 2013", 
)

# Table S2: prev studies' LENA results, along with the 59 of 61 recordings in the current dataset that are >8hrs
Bergelson_s2 <- tribble(
  ~study, ~AWC_mean, ~AWC_sd, ~CVC_mean, ~CVC_sd, ~N,
  "Gilkerson et al. 2017", 12709, 4274, 1817, 787, , # these are full-day recordings (12hrs)
  "Greenwood et al. 2011", 13142, 5562, 1714, 850, , # AWC range: 631-36563, CVC range: 11-5611
  "Soderstrom & Wittebolle", 10125, 4890, 1744, 1058, ,
  "Zimmerman et al. 2009", 12800, 4400, NA, NA, ,
  "Bergelson et al. 2018", 16510, 8718, 1432, 764, 59,
)

@@

Sperry <- Sperry %>% mutate(AllCaregiver = CDS_mean * AllCaregiverPctIncr,
                  AmbientEnv = CDS_mean * AmbientEnvPctIncr)
# "Finally, children in the poor Black Belt community heard an astounding 3,203 words per hour in their ambient environment, a 102% increase over the number of words addressed to them by their primary caregivers."
```

```{r}

```

>>>>

---
title: "Modeling Interventions"
output:
  html_document:
    df_print: paged
---

```{r, libraries, include=F}
require(tidyverse)
source("model-nonsampling.R")
```

@@

```{r, summary-tables, echo=F}
# Gilkerson et al. 2017 Table 2: standardized AWC

@@

# Bergelson et al. 2018 Table S1: 
Bergelson_s1 <- tribble(
  ~study, ~AWC_mean, ~AWC_sd, ~CDS_mean, ~CDS_sd, ~N, 
  "Schneidman & Goldin-Meadow 2013", 6254, NA, 2063, NA, 15,
  "Schneidman & Goldin-Meadow 2013", NA, NA, 2404, NA, 15,
  "Weisleder & Fernald 2013", 1084, 563, 598, 367, 29,
  "Johnson et al. 2014", 1725, NA, NA, NA, 33, # also a group of 7 with AWC ~1000/hr
  "Tamis-LeMonda et al. 2017", 2197, NA, NA, NA, 40,
  "Pancsofar & Vernon-Feagans (2006)", 2239, NA, NA, NA, 92, # averaged together mom (2559) and dad (1919)
  "Gilkerson et al. 2017", 1059, 356, NA, NA, 329,
  "Hart & Risley", NA, NA, NA, NA, NA
)

# Table S2: prev studies' LENA results, along with the 59 of 61 recordings in the current dataset that are >8hrs
Bergelson_s2 <- tribble(
  ~study, ~AWC_mean, ~AWC_sd, ~CVC_mean, ~CVC_sd, ~N,
  "Gilkerson et al. 2017", 12709, 4274, 1817, 787, 72, # these are full-day recordings (12hrs)
  "Greenwood et al. 2011", 13142, 5562, 1714, 850, 30, # mid- to upper-SES; AWC range: 631-36563, CVC range: 11-5611
  "Soderstrom & Wittebolle", 10125, 4890, 1744, 1058, 11, # mostly shorter recordings (6-8hrs)
  "Zimmerman et al. 2009", 12800, 4400, NA, NA, 275, # cross-sectional; only 71 longitudinal
  "Bergelson et al. 2018", 16510, 8718, 1432, 764, 59,
)
# from Melanie and Elika and Alex Cristia:
# N children = 264; N recordings = 851 AWC/hr across the recordings/participants:
#    0%    25%    50%      75%     100%
#  33.63  756    1087     1458    3440

@@

Sperry <- Sperry %>% mutate(AllCaregiver = CDS_mean * AllCaregiverPctIncr,
                  AmbientEnv = CDS_mean * AmbientEnvPctIncr)
# "Finally, children in the poor Black Belt community heard an astounding 3,203 words per hour in their ambient environment, a 102% increase over the number of words addressed to them by their primary caregivers."

print(Bergelson_s2)
```

Let's use the hourly rates from individual children from Sperry et al. (2018), Hart & Risley (1995), Soderstrom & Wittebolle (2013), and Weisleder & Fernald (2013) to estimate the parameters of hourly input rate distributions for both child-directed speech (CDS) and all adult speech.

```{r, cds, echo=F, warning=F, caption="Distribution of hourly child-directed speech rates heard by individual children."}
df <- read.csv("data/hourly_tokens_Sperry_HartRisley.csv")
cds_agg <- df %>% summarise(mean = mean(mother_child_tokens_hr, na.rm=T),
                           sd = sd(df$mother_child_tokens_hr, na.rm=T))
# CDS mean: 1197.5, sd: 840.2
breaks = seq(0, 3650, by = 120) 
ggplot(df, aes(x = mother_child_tokens_hr)) + 
    geom_histogram(aes(y =..density..), breaks = breaks, 
                   color = "black", fill = "gray") +
    stat_function(fun = dnorm, args = list(mean = cds_agg$mean, sd = cds_agg$sd)) +
  geom_vline(xintercept=cds_agg$mean, lty="dashed") + xlab("Child-directed Speech (Tokens/Hour)") + theme_bw() +
  annotate("text", x=cds_agg$mean +700, y=.0008, label=paste("Mean =",round(cds_agg$mean,0)))
ggsave("figures/CDS_hourly_tokens_Sperry_HartRisley_WeislederFernald.pdf", width=4.5,height=4)

# should this be a Poisson distribution? not normal
#ggplot(transform(data.frame(x=0:3650), y=dpois(x, cds_agg$mean)), aes(x, y)) + 
#    geom_bar(stat="identity")
```

```{r, awc, echo=F, warning=F, caption="Distribution of hourly adult speech rates heard by individual children."}
# all (adult) tokens Sperry + Soderstrom Wittebolle + Weisleder Fernald
awc_agg <- df %>% summarise(mean = mean(all_tokens_hr, na.rm=T),
                           sd = sd(df$all_tokens_hr, na.rm=T))
# AWC: mean: 1797.8 sd=1130.2
breaks = seq(0, 5450, by = 150) 
ggplot(df, aes(x = all_tokens_hr)) + 
    geom_histogram(aes(y =..density..), breaks = breaks, 
                   color = "black", fill = "gray") + theme_bw() +
    stat_function(fun = dnorm, args = list(mean = awc_agg$mean, sd = awc_agg$sd)) +
  geom_vline(xintercept=awc_agg$mean, lty="dashed") + xlab("Adult Word Count (Tokens/Hour)") +
  annotate("text", x=awc_agg$mean +700, y=.0008, label=paste("Mean =",round(awc_agg$mean,0)))
ggsave("figures/all_hourly_tokens_Sperry_SoderstromWittebolle_WeislederFernald.pdf", width=5,height=4)

nsubj_cds = sum(!is.na(df$mother_child_tokens_hr)) # 113 children
nsubj_awc = sum(!is.na(df$all_tokens_hr)) # 83 children
```

Based on data from `r nsubj_cds` children, the mean number of child-directed tokens heard per hour is `r round(cds_agg$mean,0)` (sd=`r round(cds_agg$sd,0)`).
Using data from `r nsubj_awc` children, the mean number of all adult tokens heard per hour is `r round(awc_agg$mean,0)` (sd=`r round(awc_agg$sd,0)`).

## References

Gilkerson, J., Richards, J. A., & Topping, K. (2017). Evaluation of a lena-based online intervention for parents of young children. Journal of Early Intervention, 1053815117718490.

Greenwood, C. R., Thiemann-Bourque, K., Walker, D., Buzhardt, J., & Gilkerson, J. (2011). Assessing children’s home language environments using automatic speech recognition technology. Communication Disorders Quarterly, 32(2), 83-92.

Soderstrom, M., & Wittebolle, K. (2013). When do caregivers talk? the influences of activity and time of day on caregiver speech and child vocalizations in two childcare environments. PloS one, 8(11), e80646.

Weisleder, A., & Fernald, A. (2013). Talking to Children Matters: Early Language Experience Strengthens Processing and Builds Vocabulary. Psychological Science, 24(11), 2143–2152.

Zimmerman, F. J., Gilkerson, J., Richards, J. A., Christakis, D. A., Xu, D., Gray, S., & Yapanel, U. (2009). Teaching by listening: The importance of adult-child conversations to language development. Pediatrics, 124(1), 342–349.
